---
title: "GDP-Aanalytics"
author: "Oral Fredric"
date: "June 21, 2016"
output: word_document
---
# Analyze the GDP and Educational data, Categorize based on Income Groups

### GDP Ranking Data

#### Year to year changes in the nominal level of output or income of an economy are affected by a combination of forces: real growth, price inflation, and exchange rates. Changes in any of the three can affect an economy's relative size and, therefore, its ranking in comparison to other economies. Of the rankings presented here, nominal GDP, perhaps the most familiar measure of aggregate economic activity, is most subject to price and exchange rate effects. Rankings are based on available data only.

### Education Statistics Data

#### The World Bank EdStats All Indicator Query holds around 3,000 internationally comparable indicators that describe education access, progression, completion, literacy, teachers, population, and expenditures. The indicators cover the education cycle from pre-primary to vocational and tertiary education.




### Download GPD Ranking Data

```{r}
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv"
gdpfile <- file.path(getwd(), "GDP.csv")
download.file(url, gdpfile)
```

### Tidying the GDP Ranking Data

```{r}
library(data.table)
GDPdata <- data.table(read.csv(gdpfile))
head(GDPdata)
GDPdata <- data.table(read.csv(gdpfile, skip=4, nrows=215))
head(GDPdata)
GDPdata <- GDPdata[X != ""]
GDPdata <- GDPdata[, list(X, X.1, X.3, X.4)]
setnames(GDPdata, c("X", "X.1", "X.3", "X.4"), c("CountryCode", "rankingGDP", "Long.Name", "gdp"))
head(GDPdata)
```


### Download Education Statistics Data

```{r}
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv"
edfile <- file.path(getwd(), "EDSTATS_Country.csv")
download.file(url, edfile)
```


### Reading Education Statistics Data

```{r}
EDdata <- data.table(read.csv(edfile))
head(EDdata)
```

### Match the Data based on CountyCode and Count How many IDs match

```{r}
GDPmergeED <- merge(GDPdata, EDdata, all=TRUE, by=c("CountryCode"))
head(GDPmergeED)
str(GDPmergeED)
sum(!is.na(unique(GDPmergeED$rankingGDP)))
```

### Sort the data frame in ascending order by GDP rank and find 13th one

```{r}
GDPmergeED[order(rankingGDP, decreasing=TRUE), list(CountryCode, Long.Name.x, Long.Name.y, rankingGDP, gdp)][13]
```

### Find the average GDP rankings for the "High income: OECD" and "High income: nonOECD" groups

```{r}
GDPmergeED[, mean(rankingGDP, na.rm=TRUE), by=Income.Group]
```

### GDP ranking into 5 separate quantile groups. Make a table versus Income.Group. How many countries are Lower middle income but among the 38 nations with highest GDP


```{r}
breaks <- quantile(GDPmergeED$rankingGDP, probs=seq(0, 1, 0.2), na.rm=TRUE)
GDPmergeED$quantileGDP <- cut(GDPmergeED$rankingGDP, breaks=breaks)
GDPmergeED[Income.Group == "Lower middle income", .N, by=c("Income.Group", "quantileGDP")]
```

### Plot the GDP for all of the countries. Use ggplot2 to color your plot by Income Group

```{r}
library(ggplot2)
ggplot(GDPmergeED,aes(CountryCode, gdp)) + geom_point(aes(color = Income.Group))
```

